Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-04-03 manuscript/Chapter2-manuscript.Rmd
update discussion

Empirical research in humans and animals shows that task proficiency often increases nonlinearly with experience and stabilizes at an expert level [reviewed in @Dukas2019]. However, we cannot derive conclusions at the individual-level from these studies either because their models assume the same increase in expertise for each individual, or because it is unclear whether their data includes repeated measures of individual success through time. In contrast, our analyses revealed that there was no population increase in expertise over time; instead, individual predators in *DBD* displayed distinct patterns of expertise acquisition. Many predators increased their success (Appendix S1: Figure S1A), but others displayed a decrease in success (Appendix S1: Figure S1B) or no change in success at all with experience. We predicted that this could occur because predators could differ in the prey that they encounter by chance, as some may be easier and others more difficult to capture. In nature, prey can use a variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. In *DBD* we showed in a previous study that faster speed and space coverage is an effective antipredator strategy used by the prey [@FraserFranco.etal2022]. Other studies also show that faster escapes are effective to escape predation [@Walker.etal2005;@Kelley.Magurran2011;Martin.etal2022]. Yet, prey speed in our current study did not shape individual differences in expertise acquisition by predators, even if predators differed in the prey that they encountered. It is however possible that other antipredator tactics mediated these individual predator responses, as we focused solely on prey speed at the group level. For example, unpredictable escapes and turning angles are important locomotor traits enabling prey to confuse human and non-human predators during the attack sequence [@Jones.etal2011;@Briffa2013;@Herbert-Read.etal2017;@Moore.etal2017;@Richardson.etal2018], which could prove to be difficult for a predator to learn and adjust to if it more often encountered prey using these tactics.

Alternatively, it is possible that individual differences in learning were also major causes, irrespective of differences in prey encounters [@Warburton2003;@morand-ferronStudyingEvolutionaryEcology2016;@Dukas2019]. *ici je pourrais plutot parler de différences d'apprentissage. Il pourrait même être + difficile d'apprendre à capturer des proies selon leur comportement car ça ne donne pas toujours autant de signaux directs que la coloration par exemple. Trouver des sources pour la coloration...ajouter que c'est irrespective of prey behaviour pour différencier par rapport à l'explication d'en haut*

 For instance, a persevering predator may increase its success much more steadily than an unmotivated predator that quickly abandons a chase. Another factor that may explain differences in the development of expertise is the time delay between hunting events. For example, a predator that played 300 matches in the span of six months might forget more critical information (e.g. prey escape patterns or muscle memory) than one that played 300 matches in the span of 6 days.

 While this has not been formally tested, a greater delay between exposure events is hypothesized to attenuate or even negate the relationship between success and experience [@Endler1991;@Wright.etal2022].

>>>>

Empirical research in humans and animals shows that task proficiency often increases nonlinearly with experience and stabilizes at an expert level [reviewed in @Dukas2019]. However, we cannot derive conclusions at the individual-level from these studies either because their models assume the same increase in expertise for each individual, or because it is unclear whether their data includes repeated measures of individual success through time. Our analyses revealed that there was no population increase in expertise over time; instead, individual predators in *DBD* displayed distinct patterns of expertise acquisition (Figure 1, Appendix S1: Figure S1A-S1B). We predicted that this would occur if, by chance, predators differed in the prey that they encountered, because some may be easier and others more difficult to capture. In nature, prey can use a variety of defenses such as physical armaments, toxins, camouflage, or antipredator behaviours to escape predation [@BrodieIII.BrodieJr.1999;@Bowen.etal2002;@brodieCoEvolutionPredatorsPrey2010;@Nomura.etal2011;@Carey.Wahl2011]. In *DBD* we previously showed that faster speed and space coverage is an effective antipredator strategy used by the prey [@FraserFranco.etal2022]. Other studies have also found that faster escapes are effective to escape predation [@Walker.etal2005;@Kelley.Magurran2011;Martin.etal2022]. Yet, despite that predators differed in the prey that they encountered, prey speed did not shape individual differences in expertise acquisition. It is however possible that other antipredator tactics mediated these individual responses. For example, it is well known in *DBD* that some prey groups use mobbing to confuse the predator or prevent it from capturing an injured individual. However, like natural predator-prey systems, direct predator-mobbing in *DBD* is dangerous and requires coordinated actions that are learned through experience [@Graw.Manser2007;@Dutour.etal2016;@Pitman.etal2017;@Dutour.etal2019;@Rampim.etal2020;@Szymkowiak2021], and was thus probably not a pervasive antipredator strategy. Other antipredator tactics known to confuse human and non-human predators during the attack sequence are unpredictable escapes and turning angles [@Jones.etal2011;@Briffa2013;@Herbert-Read.etal2017;@Moore.etal2017;@Richardson.etal2018]. While predators can adjust to unpredictable prey escapes [@Szopa-Comley.Ioannou2022], it could still be more arduous for a predator to learn and optimize its hunting success if it more often encountered prey using these tactics.

Alternatively, it is possible that individual variation in learning rates due to the time delay between hunting events drove the observed differences in expertise acquisition, irrespective of the prey groups that predators encountered. For example, in *DBD*, a predator that played 300 matches in the span of six months might forget more critical information (e.g. prey escape patterns or muscle memory) than one that played 300 matches in the span of 6 days. While this has not been formally tested, a greater delay between exposure events is hypothesized to attenuate or even negate the relationship between success and experience [@Endler1991;@Wright.etal2022]. Moreover, the predators in *DBD* may have differed in their capacity to process cues related to prey detection. For example, it is widely demonstrated that predators learn visual cues such as coloration to recognize and avoid unpalatable prey [@Skelhorn.etal2016;@Ward-Fear.etal2020;@McLellan.etal2021], or habitat characteristics to locate food patches [@Ehlinger1989;@Hughes.etal1992;@Warburton2003]. @Szopa-Comley.etal2020 also found consistent among individual variation in the capacity to detect cryptic and conspicuous prey in three-spined sticklebacks (*Gasterosteus aculeatus*). Thus, individual predators should differ in their expertise if they do not learn equally how to exploit the visual cues that facilitate prey detection [@dukasEvolutionaryBiologyAnimal2004;@Dukas2019;@Wright.etal2022].
